Evaluating the Performance and Energy Efficiency of COSMO-ART, a Fully Online Coupled Model System with Numerical Weather Forecast and Chemical Transport Models

### 2014-04-02 trunk/paper_enahpc14/Sec2:Power-measurement.tex
updates sections 1 & 2; adds file "Conclusion.tex"

\subsection{Power-Performance measurement framework}
\label{subsec:3.1}
To assess the  performance and the energy efficiency  of COSMO-ART, we
employ   a    version   of   the    integrated   framework   presented
in~\cite{energy13}  that  works in  combination  with VampirTrace  and
Vampir,   which  are   profiling/tracing   and  visualization   tools,
respectively.
%The left part of the \vref{fig:Lustre} offers a graphical representation of
%the Lustre architecture; the right depicts the tracing and profiling framework.
To  use our  approach,  COSMO-ART is  compiled  using the  VampirTrace
compiler wrappers, which automatically  instrument the Fortran code of
the  model. Next,  COSMO-ART is  run  on the  nodes, thus  dissipating
certain  amount of  power.  The  server nodes  are connected  to power
measurement  devices that  account for  the  dissipated power/consumed
energy and  send the power data  to the tracing  server.  The attached
VampirTrace \pmlib plugin employs the client API that sends start/stop
primitives in order to gather captured data by the wattmeters onto the
tracing server,  where an  instance of the  \pmlib server  is running.
Once COSMO-ART run is finished, the VampirTrace \pmlib plugin receives
the  power   data  from  the  tracing   server.   The  instrumentation
post-process  generates the  performance  trace files  and the  \pmlib
plugin inserts the power data into them.

>>>>

\subsection{Hardware description}
\label{subsec:3.1}

\subsubsection{Monch (CSCS - ETH Zurich)}
While hardware  platforms mature  and are replaced,  we have  chosen a
state-of-the-art  dual-socket Intel  Ivybridge-based  platform (called
``Monch'')  for our power  measurements, as  it is  slated to  stay in
service without upgrade for the duration of the Exa2Green project.  In
principle at least,  this architecture could be recreated  or found in
an identical configuration beyond  the lifetime of the project.  Given
that  the baseline  benchmark  can be  reproduced  within an  expected
variance   (cf  previous   section),   and  that   the  baseline   run
configuration can be  used in all future versions of  the code, a fair
comparison  will  be  made  between  the baseline  and  the  milestone
versions of COSMO-ART.

Monch, which is installed  at the Swiss National Supercomputing Center
(CSCS)  of  the  ETH  Zurich,  is a  10  rack  NEC-provided  (LX-2400)
Intel-based cluster, utilised  from people that are part  of the Swiss
Platform  for  Advanced Scientific  Computing  (PASC),  which aims  at
positioning  Swiss  computational sciences  in  the emerging  exascale
era. It is  composed of 16 administration nodes,  312 standard compute
nodes, 24 large-memory compute nodes, 24 huge-memory compute nodes and
8  servers to  provide  I/O  to Lustre.   Each  standard compute  node
comprises 2 Intel 10-core Ivy Bridge EP E5-2660v2 processors operating
at  2.2 GHz,  offering 32GB  of DDR3  1600MHz RAM,  with a  base clock
frequency of 2.2 GHz, themselves  connected by a high speed Infiniband
network,  based on Mellanox  SX6036 managed  FDR (Fourteen  Data Rate)
switches, with a 56 Gb/s speed.

A snapshot of the code, which includes, at least conceptually, all the
information needed  to reproduce the  energy-to-solution benchmarks of
COSMO-ART, was  produced and run  on a full  rack of 52  nodes without
hyperthreading (hyperthreading was attempted, but always led to higher
energy-to-solution), i.e. on a total  of 1040 cores using 20 tasks per
node.

\subsection{Power-Performance measurement framework}
\label{subsec:3.2}

\subsubsection{CSCS - ETH Zurich}
E3METER  Intelligent Power Strips  (IPS) and  Monitors (IPM)  are high
quality electricity meters which enable to monitor and optimize energy
consumption of datacenters or large facilities.  A central data logger
- the  E3METER Data Concentrator  - aggregates  all measured  data and
makes them available in various forms (SNMP, Web, etc). Data transfers
between E3METER  IPS and E3METER Concentrator  use reliable narrowband
powerline  communication (PLC)  technology which  avoids the  need for
extra cabling.  It has  one internal temperature sensor. Two dedicated
extension  ports  can be  used  to  measure  external temperature  and
humidity  through E3METER  remote sensors.  It  features non-intrusive
voltage, current and frequency measurement through current transformer
with an accuracy of 1\%.

\subsubsection{University of Hamburg}
To assess the  performance and the energy efficiency  of COSMO-ART, we
employ   a    version   of   the    integrated   framework   presented
in~\cite{energy13}  that  works in  combination  with VampirTrace  and
Vampir,   which  are   profiling/tracing   and  visualization   tools,
respectively.
%The left part of the \vref{fig:Lustre} offers a graphical representation of
%the Lustre architecture; the right depicts the tracing and profiling framework.
To  use our  approach,  COSMO-ART is  compiled  using the  VampirTrace
compiler wrappers, which automatically  instrument the Fortran code of
the  model. Next,  COSMO-ART is  run  on the  nodes, thus  dissipating
certain  amount of  power.  The  server nodes  are connected  to power
measurement  devices that  account for  the  dissipated power/consumed
energy and  send the power data  to the tracing  server.  The attached
VampirTrace \pmlib plugin employs the client API that sends start/stop
primitives in order to gather captured data by the wattmeters onto the
tracing server,  where an  instance of the  \pmlib server  is running.
Once COSMO-ART run is finished, the VampirTrace \pmlib plugin receives
the  power   data  from  the  tracing   server.   The  instrumentation
post-process  generates the  performance  trace files  and the  \pmlib
plugin inserts the power data into them.
